### 2018-04-26 stp/paper/commondefs.tex
Rename folder, fix some bib items, correct code link, update readme

\definecolor{lightgray}{rgb}{0.98, 0.98, 0.98}
%\definecolor{lightgray}{rgb}{0.83, 0.83, 0.83}
%\definecolor{orange}{HTML}{FF7F00}

% Syntax highlighting for stochastic Minizinc in listings
\lstdefinelanguage{minizinc}
{
morekeywords={
% minizinc keyworkds
any, array, assert, bool, constraint, else, endif, enum, float, forall, 
function,
if, in, include, int, list, of, op, output, minimize, maximize, par, predicate, record, set,
solve, string, test, then, tuple, type, where,
% minizinc functions
abort, abs, acosh, array_intersect, array_union,
array1d, array2d, array3d, array4d, array5d, array6d, asin, assert, atan, bool2int, card,
ceil, combinator, concat, cos, cosh, dom, dom_array, dom_size, dominance,
fix, exp, floor, index_set, index_set_1of2,
index_set_2of2, index_set_1of3, index_set_2of3, index_set_3of3, int2float, is_fixed,
join, lb, lb_array, length, let, ln, log, log2, log10, min, max, pow, product, round, set2array,
show, show_int, show_float, sin, sinh, sqrt, sum, tan, tanh, trace, ub, and ub_array},
sensitive=false, % are the keywords case sensitive
morecomment=[l][\bfseries\color{cyan!90!black}]{::},
morecomment=[l][\em\color{green!50!black}]{\%},
%morecomment=[s]{/*}{*/},
morestring=[b]",
}

% settings for listings
\lstset{ %
  backgroundcolor=\color{lightgray},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  basicstyle=\scriptsize\ttfamily,        % the size of the fonts that are used for the code
  belowskip=-.5em,
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{ForestGreen},    % comment style
%  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,                    % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\bfseries\color{blue},       % keyword style
  language=minizinc,                % the language of the code
%  morekeywords={*,...},            % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{Gray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=1,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{red},     % string literal style
  tabsize=2,                       % sets default tabsize to 2 spaces
  title=\lstname,                   % show the filename of files included with 
  %\lstinputlisting; also try caption instead of title
  % keyword highlighting
      classoffset=1, % starting new class
      morekeywords={var},
      otherkeywords={var},
      keywordstyle=\itshape\bfseries\color{red!80!black},
      classoffset=0,
      classoffset=2, % starting new class
      morekeywords={ann},
      otherkeywords={ann},
      keywordstyle=\bfseries\color{cyan!80!black},
      classoffset=0
}

%\def\mzninline{\lstinline[basicstyle=\ttfamily,annotationstyle=\normalfont]}
\def\mzninline{\verb}